---
title: "blockr.dag"
format: md
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
library(htmltools)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures",
  out.width = "100%"
)
```

# blockr.dag

<!-- badges: start -->
[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![status](https://github.com/BristolMyersSquibb/blockr.dag/actions/workflows/ci.yaml/badge.svg)](https://github.com/BristolMyersSquibb/blockr.dag/actions/workflows/ci.yaml)
[![coverage](https://codecov.io/gh/BristolMyersSquibb/blockr.dag/graph/badge.svg?token=s0zz3En4x1)](https://app.codecov.io/gh/BristolMyersSquibb/blockr.dag)
[![CRAN status](https://www.r-pkg.org/badges/version/blockr.dag)](https://CRAN.R-project.org/package=blockr.dag)
<!-- badges: end -->

An interative network library provided by g6R can be used as front-end to a blockr board using this package.

## Installation

You can install the development version of blockr.dag from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("BristolMyersSquibb/blockr.dag")
```

## Example

To start up a board with the `dag` extension, run the following code:

```{r basic-example, eval=FALSE, code=readLines("inst/examples/stacks/app.R")}
```

::: {.callout-note}
The demo below runs with shinylive. Not all feature may work as expected due to compatibility issues with webR.
:::

```{r shinylive_url, echo = FALSE, results = 'asis'}
# extract the code from knitr code chunks by ID
code <- paste0(
  c(
    "webr::unmount(\"/usr/lib/R/library/blockr.core\")",
    "webr::install(\"blockr.core\", repos = \"https://cynkra.github.io/blockr.webR/\")",
    "webr::install(\"blockr.dag\", repos = \"https://cynkra.github.io/blockr.webR/\")",
    knitr::knit_code$get("basic-example")
  ),
  collapse = "\n"
)

url <- roxy.shinylive::create_shinylive_url(code, header = FALSE)
```

```{r shinylive_iframe, echo = FALSE, eval = TRUE}
tags$iframe(
  class = "border border-5 rounded shadow-lg",
  src = url,
  style = "zoom: 0.75;",
  width = "100%",
  height = "1100px"
)
```
