---
title: "blockr.dag"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{blockr.dag}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(blockr.dag)
library(blockr.core)
library(blockr.dock)
```

## Basic Usage

### Starting a DAG Board

We create an empty board from [blockr.dock](https://github.com/BristolMyersSquibb/blockr.dock) and serve it with the `new_dag_extension()` extension:

```{r}
# Create an empty board.
serve(new_dock_board(extensions = new_dag_extension()))
```

### Pin block positions

The board is the single source of truth for the DAG: nodes, edges, combos and
their styling are always derived from it. The extension owns only
board-independent view attributes, currently node positions, passed as
`positions` (a named list keyed by block id) and persisted across
save / restore. The auto-layout currently still computes final placement at
cold start, so positions are not yet honored over it (a planned follow-up):

```{r}
serve(
  new_dock_board(
    blocks = c(
      a = new_dataset_block("iris"),
      b = new_head_block()
    ),
    links = list(from = "a", to = "b", input = "data"),
    extensions = new_dag_extension(
      positions = list(
        a = list(x = 200, y = 150),
        b = list(x = 200, y = 350)
      )
    )
  )
)
```

### Configure the layout

By default the DAG uses a built-in `antv-dagre` layout tuned to stay compact on
larger boards. Pass any [g6R](https://github.com/cynkra/g6R) layout to `layout`
to override it, e.g. lay the graph out left-to-right:

```{r}
serve(
  new_dock_board(
    blocks = c(
      a = new_dataset_block("iris"),
      b = new_head_block()
    ),
    links = list(from = "a", to = "b", input = "data"),
    extensions = new_dag_extension(
      layout = antv_dagre_layout(rankdir = "LR", nodesep = 20)
    )
  )
)
```

The chosen layout is persisted across save / restore (for layouts without
`JS()` callbacks).

## Options

blockr.dag's behaviour is tuned through a handful of R `options()`, set before `serve()`. They split into options owned by blockr.dag and options it inherits from [g6R](https://cynkra.github.io/g6R/), the engine that renders and drives the DAG.

### blockr.dag options

| Option | Default | Description |
|--------|---------|-------------|
| `blockr.dag.svg_renderer` | `FALSE` | Render the DAG with g6R's SVG renderer instead of the default canvas renderer. Canvas is the default because the SVG element reports `offsetWidth == 0`, which makes the underlying `g-lite` client/canvas coordinate scaling ignore the page zoom factor and desyncs hit-testing below 100% browser zoom (drops and port grabs silently fail). The SVG renderer keeps every element in the DOM, so it is mainly useful for `shinytest2` end-to-end tests, which opt in via `AppDriver$new(options = list(blockr.dag.svg_renderer = TRUE))`. |

### g6R options

Because blockr.dag renders through g6R, several g6R options shape DAG behaviour. Set them alongside the board:

```{r}
options(
  g6R.mode = "dev",
  g6R.directed_graph = TRUE,
  g6R.layout_on_data_change = TRUE
)

serve(new_dock_board(extensions = new_dag_extension()))
```

| Option | Default | Description |
|--------|---------|-------------|
| `g6R.mode` | `"prod"` | `"dev"` surfaces in-app notifications (e.g. why an edge drop was rejected); `"prod"` keeps them silent. Use `"dev"` while building. |
| `g6R.directed_graph` | `FALSE` | Maintain parent/child (tree) relationships as edges and nodes are added or removed through the g6R proxy. A DAG is directed, so this is typically `TRUE`; g6R also auto-enables it when nodes declare `children`. The tree bookkeeping is only kept in sync when you mutate the graph through the proxy functions, not the raw G6 JS API. |
| `g6R.layout_on_data_change` | `FALSE` | Re-run the layout whenever the graph data changes (a block or edge is added/removed), so the DAG re-arranges itself. With `FALSE`, existing nodes keep their position and only new elements are placed. |
| `g6R.preserve_elements_position` | `FALSE` | Preserve node positions across redraws instead of recomputing them. |

## User Interface Elements

### Keyboard Shortcuts

- **Shift + Drag**: Create connections between blocks.
- **Ctrl + Backspace**: Remove selected elements.
- **Alt/Option(Mac) + Drag**: Multi-select elements.

## Developer Guide

### Where does blockr.dag fit in?

In the blockr [ecosystem](https://bristolmyerssquibb.github.io/blockr/), blockr.dag is a high level extension that usually comes at the end of the stack. It depends on blockr.core for core functionalities and blockr.dock for the user interface. It also relies on external packages like g6R for rendering and interaction, as well as extra block packages like blockr.dplyr and blockr.ggplot for additional block types:

```{mermaid}
graph TD
    subgraph Core Layer
        blockr.core[blockr.core]
    end

    subgraph Extension Blocks
        blockr.dplyr[blockr.dplyr]
        blockr.ggplot[blockr.ggplot]
    end

    subgraph UI Layer
        blockr.dock[blockr.dock]
        blockr.dag[blockr.dag]
    end

    subgraph External Dependencies
        dockViewR[dockViewR]
        g6R[g6R]
    end

    blockr.core --> blockr.dock
    blockr.dock --> blockr.dag
    dockViewR --> blockr.dock
    g6R --> blockr.dag

    blockr.core --> blockr.dplyr
    blockr.core --> blockr.ggplot

    blockr.dplyr -.-> blockr.dock
    blockr.dplyr -.-> blockr.dag
    blockr.ggplot -.-> blockr.dock
    blockr.ggplot -.-> blockr.dag
```

### Maintaining the DAG state

Server-side observers handle state synchronization between the block.core __board__ and the g6R output. `update` and `board` are inherited from `blockr.core`:

`update` is composed of:

| Category | Operation | Description |
|----------|-----------|-------------|
| `blocks` | `add` | New blocks to add (blocks object) |
|          | `mod` | Modified blocks (blocks object) |
|          | `rm` | Deleted blocks (block IDS, character) |
| `stacks` | `add` | New stacks to add (stacks object) |
|          | `mod` | Modified stacks (stacks object) |
|          | `rm` | Deleted stacks (stacks IDS, character) |
| `links`  | `add` | New links to add (links object) |
|          | `mod` | Modified links (links object) |
|          | `rm` | Deleted links (links IDS, character()) |

We listen to `update()` and apply the necessary changes to the `g6R` widget via its __proxy__ functions:

```{r}
# Update observer handles board changes
update_observer <- function(update, board, proxy) {
  observeEvent(update(), {
    upd <- update()

    if (length(upd$blocks$add)) {
      add_nodes(upd$blocks$add, board$board, proxy)
    }

    if (length(upd$stacks$mod)) {
      update_combos(upd$stacks$mod, board$board, proxy)
    }

    # Handle other update types...
  })
}
```

## Extending the Package

### Custom context menu and toolbar

You can learn more in the following vignettes:

- [Context menu](https://bristolmyerssquibb.github.io/blockr.dag/articles/context-menu.html).
- [Toolbar](https://bristolmyerssquibb.github.io/blockr.dag/articles/toolbar.html).

### Extension block callback

blockr.dock exposes a callback system to extend block functionality. You can learn more in the following [vignette](https://bristolmyerssquibb.github.io/blockr.dag/articles/block-callback.html).

