---
title: "Toolbar Items"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Toolbar Items}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(blockr.dag)
```

# Toolbar Items

The toolbar provides quick access to common graph operations.

## Available Actions

- **Zoom in/out** - Navigate large workflows.
- **Auto fit** - Fit entire graph to view.
- **Layout** - Reset graph layout.
- **Add block** - Insert new processing blocks.
- **Add stack** - Create new block groupings.
- **Remove selected** - Delete selected elements.

## Creating Toolbar Items

Toolbar items are registered through the extension system:

```{r}
# Toolbar item
item <- new_toolbar_item(
  id = "custom_tool",
  icon = "icon-custom",
  js = "console.log('Custom tool clicked')",
  action = function(board, update, ...) {
    # Server-side logic
  }
)
```

Any new item has to go inside `toolbar_items.dag_extension`:

```r
toolbar_items.dag_extension <- function(x) {
  list(
    item
  )
```
